---
title: "2023_corrections"
author: "NYSDEC"
date: "`r Sys.Date()`"
output: html_document
params:
  year: 2023
  AC_match_complete: TRUE
  matched_file_name: WAVE_2023_matched.csv
---

This is a copy of Start_Here_Processing_master.Rmd
The purpose of this script is to re-run sites that were lost or duplicated. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
params$year = params$year
params$AC_match_complete = params$AC_match_complete
params$matched_file_name = params$matched_file_name
```

## Start Here
Begin by reading in your files that were downloaded from survey123

```{r read-in-files}
library(dplyr)
#read in survey data
bugs<-read.csv(paste(params$year,"/bug_repeat_1.csv",sep = ""),stringsAsFactors = FALSE)
sample<-read.csv(paste(params$year,"/WAVE_bug_id_0.csv",sep = ""),stringsAsFactors = FALSE)
field<-read.csv(paste(params$year,"/S_WSEI_SAMPLE_EVENT_INFO_0.csv",sep = ""))

#read in the master macro table for WAVE families
MACROS <- read.csv("lookuptables/Families.csv")

#limit to just the present year for all of these
bugs<-bugs %>% 
  mutate(date = format(as.Date(CreationDate,"%m/%d/%Y %H:%M:%S"))) %>% 
  filter(date >=params$year)

sample<-sample %>% 
  mutate(date = format(as.Date(Sample.Date,"%m/%d/%Y %H:%M:%S"))) %>% 
  filter(date >=params$year)

field<-field %>% 
  mutate(date = format(as.Date(Date,"%m/%d/%Y %H:%M"))) %>% 
  filter(date >=params$year)

```

```{r subset-and-compile-bug-data}
bugs.short<-bugs %>% 
  select(ParentGlobalID,Final.ID) %>% 
  distinct()

#this creates a numbered column by using the ParentGLobalID to create a sequential count by ID
bugs.short$macro_num<-ave(bugs.short$ParentGlobalID, bugs.short$ParentGlobalID, FUN=seq_along)

bugs.short<-bugs.short %>% 
  tidyr::spread(key=macro_num,value=Final.ID)

#this will be useful for the email i think

bugs.short2<-bugs %>% 
  select(ParentGlobalID,Final.ID) %>% 
  distinct()

#this is a set of various groupings to accomodate emails and such. 
```

Next combine the parent and child .csv's from teh bug ID survey
 
```{r combine-survey-results}

#wide version
combo<-merge(sample,
             bugs.short,
             by.x="GlobalID",by.y="ParentGlobalID")

#long version
combo2<-merge(sample,
              bugs.short2,
              by.x="GlobalID",by.y="ParentGlobalID")

```

Next create the assessment dataframe by using the MACRO file and the long version of the combined .csvs from the bug id survey.

```{r create-assessment-df}
#merge macros and the samples

ASSESSMENT<-merge(combo2,MACROS,by.x="Final.ID",by.y="Macroinvertebrate.Family")
#Make numeric columns
ASSESSMENT$Most <- as.numeric(ASSESSMENT$Most)
ASSESSMENT$Least <- as.numeric(ASSESSMENT$Least)
ASSESSMENT$IWLScore <- as.numeric(ASSESSMENT$IWLScore)

ASSESSMENT<-ASSESSMENT %>% 
  group_by(GlobalID,Sample.Date) %>% 
  summarize(Most = sum(Most,na.rm=TRUE),
            Least = sum(Least,na.rm=TRUE),
            IWLScore = sum(IWLScore,na.rm=TRUE)) %>% 
  ungroup()
#Convert scores to assessment categories
ASSESSMENT$assessment <- NA
ASSESSMENT$assessment <- ifelse(ASSESSMENT$Least<(-3),"Possibly Impaired",ASSESSMENT$assessment)
ASSESSMENT$assessment <- ifelse(ASSESSMENT$Most>5,"No Known Impact",ASSESSMENT$assessment)
ASSESSMENT$assessment <- ifelse(ASSESSMENT$Most<6&ASSESSMENT$Least>(-4),"No Conclusion",ASSESSMENT$assessment)
ASSESSMENT$IWL <- NA
ASSESSMENT$IWL <- ifelse(ASSESSMENT$IWLScore>23,"Excellent",ASSESSMENT$IWL)
ASSESSMENT$IWL <- ifelse(ASSESSMENT$IWLScore<24&ASSESSMENT$IWLScore>17,"Good",ASSESSMENT$IWL)
ASSESSMENT$IWL <- ifelse(ASSESSMENT$IWLScore<18&ASSESSMENT$IWLScore>11,"Fair",ASSESSMENT$IWL)
ASSESSMENT$IWL <- ifelse(ASSESSMENT$IWLScore<12,"Poor",ASSESSMENT$IWL)

#check to see that the number of original records are retained
if(nrow(ASSESSMENT) != (nrow(combo))){
 stop("Error: The number of records created in the Assessment don't match the number of original records!")
}


```

 Then merge the assessment file back with the original long results
```{r merge-assessment-with-original}
#merge back with original to get the assessments
combo3<-combo2 %>% #create a shortened file of the long version
  select(!c(Final.ID,"ObjectID" )) %>% 
  distinct()

final<-merge(combo3,ASSESSMENT,by="GlobalID",all.x = TRUE)

combo4<-merge(ASSESSMENT,bugs.short,
              by.x="GlobalID",
              by.y="ParentGlobalID",all.x = TRUE) #not sure this one is helpful


#check to see that the number of original records are retained
if(nrow(final) != (nrow(combo))){
 stop("Error: The number of records created in the Assessment don't match the number of original records!")
}
```

One more merge for all the data to get the bugs that were in there
```{r}
#merge with the wide bugs to get what was in there
final<-merge(final,bugs.short,by.x = "GlobalID",by.y="ParentGlobalID")

```

Create file to match up accession numbers with field records to be able to email everyone their results.

```{r write-AC-match-Table}
# write the tables 
if(!params$AC_match_complete){#write.table(final,file="2022/Assessment_check.csv",sep=",",row.names=FALSE)
write.table(field, file =paste(params$year,
                               "/field_check_match_ACnum.csv",
                               sep = ""),
            sep = ",",row.names = FALSE)
  stop("Please use the created file to match up AC numbers, emails and people to be able to email results")
}

```

Now, we read in the matched file, this invariably does not have all of the samples, some people either don't provide paper copies and/or don't fill out the field survey.

```{r read-in-matched-file}
#read matched file back in
matched<-read.csv(here::here(paste(params$year,
                      params$matched_file_name,
                      sep = "/")))

final.short<-final %>% 
  select(Please.enter.the.Sample.Accession.Number,GlobalID) %>% 
  distinct()

combo5<-merge(combo4,final.short,by="GlobalID",all.x = TRUE)
combo6<-merge(combo5,matched,
              by.x="Please.enter.the.Sample.Accession.Number",
              by.y="AC_number",
              all.x = TRUE)

# combo6<-combo6 %>% #this is here in case you get weird columns
#   select(-c("ï..ObjectID","GlobalID.y")) %>% 
#   distinct()


```

Here is some more processing to take care of the matching for the emails

```{r}
bugs.short3<-bugs %>% 
  select(ParentGlobalID,Final.ID) %>% 
  distinct() %>% 
  group_by(ParentGlobalID) %>% 
   summarise(final_bug_id = toString(Final.ID)) #this is creating one column with all the bugs in it

#make a shortened version of the combo
combo6<-combo6 %>% 
  select(-c(9:28)) #check this, i'm trying to take out the raw bugs bc, i made the column that combined them

combo7<-merge(combo6, #this one is combining the new combo file with the raw bugs that were made into one column
              bugs.short3, 
              by.x="GlobalID.x",
              by.y="ParentGlobalID",
              all.x = TRUE)


write.csv(combo7,file = here::here(paste(params$year,
                              "/all_with_assessment_for_emails.csv",
                              sep = ""),
                              row.names = FALSE))

```

Great, now you can use that file to email the recipients that provided emails.


